\section{Design}
\label{sec:mercury-design}

In this section we describe the design of Mercury. This section will be primarily of interest to
developers who are working on the implementation of Mercury. Developers who are using Mercury to
generate API packages for CubedOS modules should refer to the tutorial in
Section~\ref{sec:mercury-tutorial}.

\subsection{Mercury Pipeline}

\begin{itemize}
\item \textbf{Scala}. Mercury is written in Scala, a functional/OOP hybrid language for the Java
Virtual Machine (JVM). Mercury capitalizes on the Scala/Java ecosystem in terms of tooling (such
as SBT) and library support.

\item \textbf{Grammar}. MXDL is described as an ANTLR4 grammar that is then processed by ANTLR
to generate the lexical analyzer and parser (in Java).

\item \textbf{Lexer}. Mercury's lexical analyzer is ANTLR generated. It is fairly standard in
terms of its handling of whitespace, identifiers, numeric literals, etc.

\item \textbf{Parser}. The parser is also ANTLR generated, and Mercury uses ANTLR generated
classes for traversing the parse tree produced by the parser. Syntax errors are detected by the
ANTLR generated code.

\item \textbf{Semantic Analyzer}. After parsing, the semantic analyzer walks over the parse tree
searching for type errors and other violations of MXDR's rules. After this stage of processing,
the input MXDR is considered valid and code can be generated from it.

\item \textbf{Code Generation}. The code generator takes the validated MXDR and generates the
Ada package that defines the API to the module. Two files are produced: the package
specification and the package body. Code generation entails walking over the parse tree again
and emitting Ada code based on the nodes encountered.
\end{itemize}

Since approximately half of the files in this Mercury compiler pipeline are auto-generated by
the ANTLR tool, the three distinct areas of focus at the code level are the grammar, the
semantic analyzer, and the code generation. SPARK skeletons are filled in by the code generation
phase after the previous compiler phases check the soundness and validity of the modified XDR.
